---
title: "Week 14 - Gettin' Extra Shiny" 
subtitle: ""
author: "Alex Lishinski"
date: "November 16, 2021"
output:
  xaringan::moon_reader:
    css: [theme.css]
    lib_dir: libs
    nature:
      highlightStyle: github
      highlightLines: true
      countIncrementalSlides: false
---

```{r, include = FALSE}
knitr::opts_chunk$set(echo = TRUE, fig.align = "center", message = FALSE)
```

# Welcome!

Welcome to *week 14*!

**Record the meeting** 

```{r, include = TRUE, echo = FALSE}
library(here)
library(tidyverse)
library(visdat)
library(mice)
```

---

## A recap of last week

- Different behaviors responding to mouse inputs in shiny 
- Shiny notifications and modal dialog boxes
- Dynamic UI with `renderUI` and `uiOutput`

---

### Topics for today

**Record the meeting** 

- A. Davery's feedbackWarning issue
- B. The issue with modal dialogs example from last time
- C. `reactiveVal` to create reactive values

---

### A. Davery's feedbackWarning issue

- Davery brought up an issue on slack with trying to use feedbackWarning
- feedbackWarning responds to "events", which can be reactive values and functions
- How you align the event with the warning will change when and how the warning shows up

---

### B. The issue with modal dialogs example from last time

- You may have noticed an issue with the modal dialog last time
- If we want the modal dialog to control downstream computations we need to make that explicit
- The action buttons in the modal dialog work just like any other for controlling other output

---

### C. `reactiveVal` to create reactive values

- The `reactiveVal` function allows you to create single reactive values 
- These values can respond to other events
- They create reactive dependencies just like other reactives so they will update accordingly when they need to


---

## Logistics

**This week**

* Homework 10: Available 11/18; **Due by 11/30 **


  
---

# Final project recordings/presentations

- Presentations in class 11/30: ~15 minutes, slides + showing products

- Strict deadline of 11:59 pm ET on 12/7

