---
title: "Week 9 - Models and Shiny" 
subtitle: ""
author: "Alex Lishinski"
date: "October 12, 2021"
output:
  xaringan::moon_reader:
    css: [theme.css]
    lib_dir: libs
    nature:
      highlightStyle: github
      highlightLines: true
      countIncrementalSlides: false
---

```{r, include = FALSE}
knitr::opts_chunk$set(echo = TRUE, fig.align = "center", message = FALSE)
```

# Welcome!

Welcome to *week 9*!

```{r echo = FALSE, out.width = "70%"}
```

**Record the meeting** 

```{r, include = TRUE, echo = FALSE}
library(here)
library(tidyverse)
```

---

### This week's topics

**Overview**

A. Final project presentations

B. A buffet of models

C. Intro to Shiny

---

### A. Final project presentations

* Add your ideas to the final project brainstorm! 

  * https://docs.google.com/presentation/d/1etv_oXeSjn8YdgZFHzPrAM4RRESXiEDFkdmodnNW1Mo/edit?usp=sharing

* Thursday, each of you will briefly (1 min. or so) present on your final project idea

* Next week, incorporate feedback present more fleshed out plan for further feedback(3-5 mins)

---

### B. A buffet of models

There are a number of ways to understand variables about which you have data and the
relationships between them.

One way is to create a **model**, a simplified _representation_ of your data that
can be informative to you (and others) about your data - and, maybe, what your data
represents. 

From this broad definition, models can take many different forms:

- A sample statistic (e.g., a _mean_ of a variable)
- A relationship describing how two variables co-vary (e.g., a bivariate _correlation_)
- A linear regression model
- . . . (what models are common in your field?)

---

### B. A buffet of models

One of the benefits of modeling your data within R is that many R packages share a 
common modeling syntax, or interface: the formula syntax.

This code represents the regression of `hp` upon `mpg`:

```{r, eval = FALSE}
mpg ~ hp
```

This code often corresponds to the underlying mathematical/statistical equation:

```{r, echo = FALSE}
m1 <- lm(mpg ~ hp, data = mtcars)
equatiomatic::extract_eq(m1)
```

---

### B. A buffet of models

Today, we'll focus on the linear regression model, but will also touch on the following:

- *t*-test
- ANOVA
- generalized linear model (i.e., Poisson or Logistic Regression)
- multi-level (or hierarchical linear) model

---

### B. A buffet of models

There is a _lot_ we can do with a linear regression model!

```{r, warning = FALSE, message = FALSE}
d <- read_csv("https://raw.githubusercontent.com/data-edu/dataedu/master/data-raw/wt01_online-science-motivation/processed/sci_mo_processed.csv")

d
```

---

### B. A buffet of models

Estimating a model; seeing the result:

```{r, warning = FALSE, message = FALSE}
lm(FinalGradeCEMS ~ TimeSpent_hours, data = d)
```
---

### B. A buffet of models

Saving the output to an _object_ and printing a summary of the results

```{r, warning = FALSE, message = FALSE}
m1 <- lm(FinalGradeCEMS ~ TimeSpent_hours, data = d)
summary(m1)
```

---

### B. A buffet of models

Making the model more complex - a multiple regression

```{r, warning = FALSE, message = FALSE}
m2 <- lm(FinalGradeCEMS ~ TimeSpent_hours + int + Gender, data = d)
summary(m2)
```

---

### B. A buffet of models

Adding an interaction 

```{r, warning = FALSE, message = FALSE}
m3 <- lm(FinalGradeCEMS ~ TimeSpent_hours + int*Gender, data = d)
summary(m3)
```

---

### B. A buffet of models

*t*-test

```{r}
m_t_test <- t.test(FinalGradeCEMS ~ Gender, data = d)
m_t_test
```

---

### B. A buffet of models

ANOVA

```{r}
m_anova <- aov(FinalGradeCEMS ~ subject, data = d)
m_anova
```

---

### B. A buffet of models

Multi-level model

```{r}
library(lme4)
m5 <- lmer(FinalGradeCEMS ~ TimeSpent_hours + int*Gender + (1|course_id), data = d)
summary(m5)
```

---

### C. Intro to Shiny

Shiny is a framework for making interactive data tools in R

Shiny projects are web apps that are made for viewing in a browser

Simple to get started, but unlimited customizability

---

### C. Intro to Shiny

Shiny apps are meant to be interactive - user changes the input, output changes

Output can take whatever form you want - plots are one good option

Creating the app involves 2 main parts:

- UI: how the app looks
- Server: how the app behaves

Shiny lets you define these 2 parts together in one file or separately

---

### C. Intro to Shiny

Minimal example of app code

```{r, eval = FALSE}
library(shiny)

ui <- fluidPage(
  "Hello, world!"
)

server <- function(input, output, session) {
}

shinyApp(ui, server)
```

---

### C. Intro to Shiny

Different layout functions let you set up basic template of UI

e.g. sidebarLayout()

Panel functions let you define the layout of different areas of the UI based on the layout

e.g. sidebarPanel(), mainPanel()

---

### C. Intro to Shiny

Within UI panels you can use various inputs from users to control the output

e.g.:

* selectInput() - drop down menu
* sliderInput() - numerical slider
* numericInput() - numerical text box

Important that you align these correctly!

---

# Logistics

**This week**

* Homework 8: Available Thursday**

* Readings:
  * https://datascienceineducation.com/c09.html
  * https://moderndive.com/5-regression.html
---

# Final Project

* [Final project]()
  * Flesh out final project idea based upon feedback (this forthcoming week)
    * Present and receive more feedback next week

---

### Wrapping up

On Slack:

- What is one thing you learned today?
- What is something you want to learn more about? 
- Share your feelings in GIF form!
