---
title: "Week 15 - Some Final Packages" 
subtitle: ""
author: "Alex Lishinski"
date: "November 23, 2021"
output:
  xaringan::moon_reader:
    css: [theme.css]
    lib_dir: libs
    nature:
      highlightStyle: github
      highlightLines: true
      countIncrementalSlides: false
---

```{r, include = FALSE}
knitr::opts_chunk$set(echo = TRUE, fig.align = "center", message = FALSE)
```

# Welcome!

Welcome to *week 15*!

**Record the meeting** 

```{r, include = TRUE, echo = FALSE}
library(here)
library(tidyverse)

library(plotly)
library(leaflet)
```

---

### Topics for today

**Record the meeting** 

- A. leaflet package
- B. plotly package

---

### A. leaflet package

- The leaflet package is a package for generating interactive map visualizations
- The maps generated using this package use map tiles from OpenStreetMap, which looks a lot like the maps from google maps
- There are many parameters you can add to your maps including different layers, markers, etc.

---

### A. leaflet package

- The interface to leaflet is fairly straightforward
- leaflet maps begin with a call of the leaflet() function
- other map components are then added on in layers, similar to ggplot
- unlike ggplot, these layers come from piping `%>%`, rather than `+`

---

### A. leaflet package

- The `leaflet()` function provides the blank (interactive) map canvas

```{r}
m <- leaflet()

m
```


---

### A. leaflet package

- The `leaflet()` function provides the blank (interactive) map canvas
- The `addTiles()` function adds your map tiles

```{r}
m <- leaflet() %>%
      addTiles()
m
```

---

### A. leaflet package

- The `leaflet()` function provides the blank (interactive) map canvas
- The `addTiles()` function adds your map tiles
- The `addMarkers()` function lets you add marker points to your map

```{r}
amy <- filter(storms, name == "Amy")

m <- leaflet() %>% 
  addTiles() %>%
  addMarkers(lat = c(min(amy$lat), max(amy$lat)), lng = c(min(amy$long), max(amy$long)))
m
```

---

### A. leaflet package

- leaflet visualizations can be incorporated into shiny
- just like the other types of output we have used with shiny, there are functions for this type
- the renderLeaflet() and leafletOutput() functions will do what you expect

---

### A. leaflet package

- There are many other things that leaflet can do
- You can add different map layers like topographical map
- You can add shapes and colors to highlight things
- Documentation and Examples here: https://rstudio.github.io/leaflet/

---

### B. plotly package

- Plotly is a package for doing interactive data visualizations
- Uses a fairly different sort of syntax from ggplot, but still rather straightforward
- Adds a number of interactive tools to the sorts of data visualizations you'd create with ggplot
- basic functionality covers some of the use cases from shiny that we've covered

---

### B. plotly package

- Plotly plots are based around the `plot_ly()` function
- Follows a roughly similar structure to ggplot
- Data argument (optional)
- Attributes (different for each plot, analogous to aes())
- type (analogous than the various `geom_` functions)
- mode (covers some differences between `geom_` functions)

---

### B. plotly package

```{r}
plot_ly (
x = c( 1, 2, 3 ),
y = c( 5, 6, 7),
type = 'scatter',
mode = 'lines')
```

---

### B. plotly package

```{r}
plot_ly (
x = c( 1, 2, 3 ),
y = c( 5, 6, 7 ),
type = 'scatter' ,
mode = 'markers' )
```

---

### B. plotly package

```{r}
plot_ly(
x = c( 1, 2, 3 ),
y = c( 5, 6, 7 ),
type = 'scatter',
mode = 'markers' ,
size = c( 1, 5, 10 ),
marker = list(
color = c( 'red', 'blue' ,
'green' )))
```


---

### B. plotly package

```{r}
plot_ly(
x = c( 1, 2, 3 ),
y = c( 5, 6, 7 ),
type = 'scatter' ,
mode = 'lines' ,
fill = 'tozeroy' )
```


---

### B. plotly package

- like leaflet, plotly visualizations can be combined with shiny rather nicely
- Also similarly, there are corresponding functions for render and output
- renderPlotly() and plotlyOutput() will do the hard work for you
- Documentation can be found here: https://plotly.com/r/plotly-fundamentals/
- Cheatsheet can be found here: https://images.plot.ly/plotly-documentation/images/r_cheat_sheet.pdf
- Book: https://plotly-r.com/index.html

---

## Logistics

* Homework 10:  **Due by 11/30 **

## Final project recordings/presentations

- Presentations in class 11/30: ~15 minutes, slides + showing products

- Strict deadline of 11:59 pm ET on 12/7
  
---
