---
title: "Week 13 - Gettin' Extra Shiny" 
subtitle: ""
author: "Alex Lishinski"
date: "November 9, 2021"
output:
  xaringan::moon_reader:
    css: [theme.css]
    lib_dir: libs
    nature:
      highlightStyle: github
      highlightLines: true
      countIncrementalSlides: false
---

```{r, include = FALSE}
knitr::opts_chunk$set(echo = TRUE, fig.align = "center", message = FALSE)
```

# Welcome!

Welcome to *week 13*!

**Record the meeting** 

```{r, include = TRUE, echo = FALSE}
library(here)
library(tidyverse)
library(visdat)
library(mice)
```

---

## A recap of last week

- The map() functions can help you to read in multiple files

- Cases with missing values can still be used with multiple imputation

---

### Topics for today

**Record the meeting** 

- A. More Shiny mouse behaviors
- B. More Shiny validation

---

### A. Shiny mouse behaviors

- Your app can respond to mouse inputs
- We saw this previously in defining the click behavior in a `plotOutput`
- The click behavior can then be connected to its own output
- There are more ways that we can define this behavior

---

### A. Shiny mouse behaviors

- In our previous use of `click =` we created an input based on clicks
- Our previous version gave us the coordinates on the plot wherever we clicked
- We can also define the behavior to give us information based on the points 
- `nearPoints()` gives us points within a certain distance of the click
- `brush()` lets us define the area of interest

---
  
### A. Shiny mouse behaviors

- In addition to clicks we can create output based on other mouse actions, like hover
- This works rather similarly to defining click behaviors, just use different arguments

---

### B. Shiny validation

- Previously we talked about `shinyFeedback` and the `feedbackWarning` and similar functions
- We also talked about the `req()` function which can be used to stop execution of code until a required condition is met
- Another tool for pre-empting possible error messages is the function `validate()`
- `validate()` can let us check for complex conditions

---

### B. Shiny validation

- Shiny also lets you give notifications to the user based on a reactive
- The basic way to fit them is to pair them with `observeEvent()`
- Notifications allow you to change the duration and type
- `Sys.sleep()` function lets you add delay between multiple notifications
- These can also respond to an ongoing task

---

### B. Shiny validation

- You can also request that a user confirm an action with a dialog box
- You can create them with `modalDialog()` 
- `showModal()` can create these dialogs and `removeModal()` gets rid of them
- These are also a use case for `observeEvent()`

---

## Logistics

**This week**

* Homework 10: Available 11/18; **Due by 11/30 **

* Reading:
  * Big Magic With R: Creating Learning Beyond Fear: https://alison.rbind.io/talk/2018-cascadia-bigmagic/
  
---

# Final project recordings/presentations

- Presentations in class 11/30: ~15 minutes, slides + showing products

- Strict deadline of 11:59 pm ET on 12/7

