---
title: "Week 10 - Modeling Data and More Shiny" 
subtitle: ""
author: "Alex Lishinski"
date: "October 19, 2021"
output:
  xaringan::moon_reader:
    css: [theme.css]
    lib_dir: libs
    nature:
      highlightStyle: github
      highlightLines: true
      countIncrementalSlides: false
---

```{r, include = FALSE}
knitr::opts_chunk$set(echo = TRUE, fig.align = "center", message = FALSE)
```

# Welcome!

Welcome to *week 11*!

```{r echo = FALSE, out.width = "70%"}
```

**Record the meeting** 

```{r, include = TRUE, echo = FALSE}
library(here)
library(tidyverse)
```

---

### Topics for today

**Record the meeting** 

Modeling continued

A. Recap from the buffet and further expansion

B. Model outputs and summaries

C. Model helpers - tests, diagnostics, and model components

D. Shiny continued

---

### A. Buffet of models recap

What a model is:

**model**: a simplified _representation_ of your data that
can be informative to you (and others) about your data - and, maybe, what your data
represents. 

From this broad definition, models can take many different forms:

- A sample statistic (e.g., a _mean_ of a variable)
- A relationship describing how two variables co-vary (e.g., a bivariate _correlation_)
- A linear regression model

---

### A. Buffet of models recap

How do we represent model equations in R? Many R packages share a common modeling syntax, or interface: the formula syntax.

This code represents the regression of `hp` upon `mpg`:

```{r, eval = FALSE}
mpg ~ hp
```

Then there are other formula operators:

```{r, eval = FALSE}
# additional independent variables
mpg ~ hp + disp
# interactions with main effects
mpg ~ hp + disp + hp*disp
# interactions without main effects
mpg ~ hp + disp + hp:disp
# All remaining variables
mpg ~ .
```

---

### A. Buffet of models recap

There are a number of helper functions that work with `lm` and other models

And there's some more advanced formula tricks too

```{r}
# Polynomial Regression
y ~ x + I(x^2) + I(x^3)

# Factorial ANOVA
y ~ (a*b*c)^2

# Variable transformations
Sepal.Width ~ Petal.Width + log(Petal.Length) + Species
```

---

### A. Buffet of models recap

The lm() function is the base case of using these model formulas

```{r, warning = FALSE, message = FALSE, include = FALSE}
d <- read_csv("https://raw.githubusercontent.com/data-edu/dataedu/master/data-raw/wt01_online-science-motivation/processed/sci_mo_processed.csv")
```

Estimating a model; seeing the result:

```{r, warning = FALSE, message = FALSE}
lm(FinalGradeCEMS ~ TimeSpent_hours, data = d)
```

---

### A. Buffet of models recap

The lm() function is the base case of using these model formulas

Saving the output to an _object_ and printing a summary of the results

```{r, warning = FALSE, message = FALSE}
m1 <- lm(FinalGradeCEMS ~ TimeSpent_hours, data = d)
summary(m1)
```

---

### A. Buffet of models recap

The lm() function is the base case of using these model formulas

Making the model more complex - a multiple regression

```{r, warning = FALSE, message = FALSE}
m2 <- lm(FinalGradeCEMS ~ TimeSpent_hours + int + Gender, data = d)
summary(m2)
```

---

### A. Buffet of models recap

But other functions and packages use this as well

t-test

```{r}
m_t_test <- t.test(FinalGradeCEMS ~ Gender, data = d)
m_t_test
```

ANOVA

```{r}
m_anova <- aov(FinalGradeCEMS ~ subject, data = d)
m_anova
```

---

### A. Buffet of models recap

Some packages add to this syntax too -- Multi-level model

```{r}
library(lme4)
m5 <- lmer(FinalGradeCEMS ~ TimeSpent_hours + int*Gender + (1|course_id), data = d)
summary(m5)
```

---

### A. Buffet of models recap

Weights using `lm()`

```{r}
# Adding weights
wts1 <- rnorm(n = nrow(d), mean = 0, sd = 1)
wts2 <- rnorm(n = nrow(d), mean = 0, sd = 1)

wts <- abs(wts1/wts2)

m1_wt <- lm(FinalGradeCEMS ~ TimeSpent_hours, data = d, weights = wts)
summary(m1_wt)
```


---

### A. Buffet of models recap

The `glm()` function - the way to fit other types of regression models

`glm()` lets you specify distribution families for different kinds of outcomes:

* "gaussian" = standard regression for continuous dependent variables (default)
* "binomial" = binary (0/1) outcome variables
* "poisson" = count outcome variables
* several others are available

---

### A. Buffet of models recap

The base case with default options will be same as `lm()`

```{r}
glm1 <- glm(FinalGradeCEMS ~ TimeSpent_hours, data = d)
summary(glm1)
```

---

### A. Buffet of models recap

The `family` argument to `glm()` helps you change the kind of model being estimated.

```{r}
d <- d %>%
  mutate(Points_Earned_Bin = ifelse(d$Points_Earned > mean(d$Points_Earned, na.rm = T), 1, 0))

glm1 <- glm(Points_Earned_Bin ~ TimeSpent_hours, data = d, family = "binomial")
summary(glm1)
```

---

### B. Model outputs and summaries

We've previously discussed making better looking output tables in .Rmd documents

```{r}
library(sjPlot)

tab_model(m1)
```

---

### B. Model outputs and summaries

Another package makes it easier to pull out model estimates in an easier format to use for further operations: `broom`

`tidy()` Makes `lm()` coefficient output into a tidy data frame format, which you can then use all of your tidyverse tools on

`glance()` works similarly but with the whole model diagnostic statistics such as $R^2$

```{r}
library(broom)

# Model coefficents
tidy(m1)

# whole model stats
glance(m1)
```


---

### B. Model outputs and summaries

`broom` can also help you grab other diagnostics from the model with `augment()`

```{r}
# Data level additional model generated values
head(augment(m1))
```

---

### B. Model outputs and summaries

`broom` can also help you grab other diagnostics from the model


```{r, out.width="50%"}
f <- augment(m1)

ggplot(f) +
  geom_point(aes(x=.fitted, y=.resid))
```

---

### B. Model outputs and summaries

The `car` package has plotting and testing functions that can help you evaluate models

```{r, out.width="50%"}
library(car)

avPlots(m2)
```

---

### B. Model outputs and summaries

The `car` package has plotting and testing functions that can help you evaluate models

```{r, out.width="50%"}
residualPlots(m2)
```

---

### B. Model outputs and summaries

The `report` package can generate written summaries of the results of your models

```{r}
#devtools::install_github("https://github.com/easystats/report")
library(report)

report(m1)
```

---

### B. Model outputs and summaries

Components of an `lm()` object, some of which you can get with helper functions

```{r}
m1$coefficients
head(m1$residuals)
```

---

### C. Model helpers - tests, diagnostics, etc.

There are a number of helper functions that work with `lm` and other models

```{r}
# model coefficients
coef(m1)

# model residuals
head(residuals(m1))

# regression fitted values
head(fitted(m1))

# Get the data used to fit the model
head(model.frame(m1))
```

---

### C. Model helpers - tests, diagnostics, etc.

There are a number of helper functions that work with `lm` and other models - How do we know which ones?

```{r}
library(sloop)
s3_methods_class("lm")

confint(m1)
vcov(m1)
```

---

### C. Model helpers - tests, diagnostics, etc.

The `lmSupport` package provides number of additional tools for regression models

```{r}
library(lmSupport)

m1 <- lm(FinalGradeCEMS ~ TimeSpent_hours, data = d)
m2 <- lm(FinalGradeCEMS ~ TimeSpent_hours + int + Gender, data = d)

# A function to check regression model assumptions
modelAssumptions(m1)
```

---

### C. Model helpers - tests, diagnostics, etc.

The `lmSupport` package provides number of additional tools for regression models

```{r}
m2 <- lm(FinalGradeCEMS ~ TimeSpent_hours + int + Gender, data = d)
m1 <- lm(FinalGradeCEMS ~ TimeSpent_hours, data = model.frame(m2$model))

# F Test difference between 2 models
modelCompare(m1, m2)
```

---

### D. Shiny Continued

*Storms App Additional Features*

Fixing the multiple hurricanes issue 

Adding a zoom slider

Adding Color

Data table under plot

---

### Addendum: Other modeling packages to be aware of

* `lme4` and `nlme`: hierarchical linear models (aka multilevel models)
* `lavaan`: structural equation models
* `MASS`: Robust regression
* `caret`: Lots of different classification and regression models for machine learning applications
* `parsnip`: (tidyverse) Unified interface to many different models, largely machine learning type
* Time series: https://cran.r-project.org/web/views/TimeSeries.html

Other types of models that may be more applicable to your field can also be found in the CRAN task views:
https://cran.r-project.org/web/views/

Tutorials for many different varieties of Regression can be found here: 
https://stats.idre.ucla.edu/other/dae/

Other packages that can help you with your models
https://easystats.github.io/easystats/
e.g. the `performance` package does model assumption checking and model comparison

---

# Logistics

**This week**

* Homework 8: Available tomorrow, Due next Tuesday 

* Reading: * https://www.tmwr.org/base-r.html#formula

---

# Assignment updates

* [Final project](https://f21-intro-to-data-sci-methods-in-ed.netlify.app/assignment-final-project/)
  * NEXT Thursday - Give updated version of your idea 

---

### Wrapping up

On Slack channel:

- What is one thing you learned today?
- What is something you want to learn more about? 
- Share your feelings in GIF form!
