---
title: "Week 11 - Functions and Iteration" 
subtitle: ""
author: "Alex Lishinski"
date: "October 25, 2021"
output:
  xaringan::moon_reader:
    css: [theme.css]
    lib_dir: libs
    nature:
      highlightStyle: github
      highlightLines: true
      countIncrementalSlides: false
---

```{r, include = FALSE}
knitr::opts_chunk$set(echo = TRUE, fig.align = "center", message = FALSE)
```

# Welcome!

Welcome to *week 11*!

```{r echo = FALSE, out.width = "70%"}
```

**Record the meeting** 

```{r, include = TRUE, echo = FALSE}
library(here)
library(tidyverse)
```

---

## A recap of last week (on modeling)

- A model is a simplification (and a summary) of your way
- A common type of model is a regression model, aka a linear regression model (or a linear model)


This code represents the regression of `hp` upon `mpg`:

```{r, eval = FALSE}
lm(mpg ~ hp, data = my_data)
```

This code often corresponds to the underlying mathematical/statistical equation:

```{r, echo = FALSE}
m1 <- lm(mpg ~ hp, data = mtcars)
equatiomatic::extract_eq(m1)
```

---

## A recap of last week (on modeling)

- A linear model can help you learn how one or more _independent variables_ (or, _x_ variables) relate to one _dependent variable_ (a _y_ variable)
- It is common to specify a number of models, and to compare them (commonly based on how much of the variation in the dependent variable is accounted for by relationships with the _x_ variable(s))
- the `lm()` function is a great tool for specifying relatively simple linear regression models, but it can be extended in a number of powerful ways
- R has a number of tools for interpreting and presenting the output of models

---

## Checking-in on final projects

- We will hear more details about your plans on Thursday
- And, generally, please consider a) the scope of your project and b) the level of detail in your plans
- I want you to do a small(er) number of things well 

---

### Topics for today

**Record the meeting** 

A. Functions  
B. Iteration (or, _applying_ functions)  
C. More shiny - Reactivity 

---

### A. Functions

A function is a collection of code that:

- takes one or more inputs (most commonly in R, data!)
- and produces one or more forms of output (often, your data---transformed!)

You _already_ use functions all of the time:

```{r}
mpg
```

---

### A. Functions

A function is a collection of code that:

- takes one or more inputs (most commonly in R, data!)
- and produces one or more forms of output (often, your data---transformed!)

You _already_ use functions all of the time:

```{r}
print(mpg)
```

---

### A. Functions

**`select()` is a function**

```{r}
mpg %>% 
  select(model, displ, cyl)
```

---

### A. Functions

**`select()` and `filter()` are functions**

```{r}
mpg %>% 
  select(model, displ, cyl) %>% 
  filter(cyl == 6)
```

---

### A. Functions

There are all kinds of functions:

```{r}
mpg %>% 
  summarize(mean_cty = mean(cty),
            mean_hwy = mean(hwy))
```

```{r}
ten_mpg_values <- c(13, 63, 23, 43, 23, 33, 42, 34, 21, 23)
mean(ten_mpg_values)
```

---

### A. Functions

You can write your own functions!

Suppose we want to write a function to _standardize_ the `cty` variable to have _M_ = 0 and _SD_ = 1;
presently, its *M* and *SD* are:

```{r}
mpg %>% 
  summarize(mean_cty = mean(cty),
            sd_cty = sd(cty))
```
---

### A. Functions

(In this case, there _is_ an existing function, but it has some quirks, including, importantly, returning a _matrix_, rather than a vector/"a column")

So, let's write our own. Here is a template, with two "blanks" represented by "___":

```{r, eval = FALSE}
standardize_a_variable <- function(___) {
  ___
}
```

---

### A. Functions

**A not very useful function!**

```{r}
standardize_variable <- function(variable) {
  variable
}
```

```{r}
mpg %>% 
  mutate(cty_std = standardize_variable(cty)) %>% 
  summarize(mean_cty_std = mean(cty_std),
            sd_cty_std = sd(cty_std))
```

---

### A. Functions

**A useful `standardize_variable`**

```{r}
standardize_variable <- function(variable) {
  variable <- variable - mean(variable) # transforms variable to have M = 0
  variable <- variable / sd(variable) # transforms variable to have SD = 1
  variable
}
```

```{r}
mpg %>% 
  mutate(cty_std = standardize_variable(cty)) %>% 
  summarize(mean_cty_std = mean(cty_std),
            sd_cty_std = sd(cty_std))
```

---

### A. Functions

**Let's double-check that the mean is practically equal to 0** (by rounding to precision to three decimal points)

```{r}
mpg %>% 
  mutate(cty_std = standardize_variable(cty)) %>% 
  summarize(mean_cty_std = mean(cty_std),
            sd_cty_std = sd(cty_std)) %>% 
  mutate(mean_cty_std_rounded = round(mean_cty_std, 3))
```

---

### A. Functions

We can add an argument to a function to modify how it works.

```{r}
standardize_variable <- function(variable, remove_na_values = FALSE) {
  variable <- variable - mean(variable, na.rm = remove_na_values)
  variable <- variable / sd(variable, na.rm = remove_na_values)
  variable
}

```

---

### A. Functions

What if we wanted to scale the variables cty, hwy, and cyl? Imagine that cyl has a missing (`NA`) value.

```{r}
mpg %>% 
  mutate(cty_std = standardize_variable(cty),
         hwy_std = standardize_variable(hwy),
         cyl_std = standardize_variable(cyl, remove_na_values = TRUE)) %>% 
  select(cty_std, hwy_std, cyl_std, everything()) # bring the new variables to the beginning of the data frame
```

---

### A. Functions

In summary:

- You already use functions all of the time
- You can write your own functions that take one or more types of input and return output
- The functions that you are using within R were written by other people!
- https://github.com/tidyverse/dplyr/blob/master/R/select.R
- https://github.com/SurajGupta/r-source/blob/master/src/library/stats/R/median.R
- https://github.com/jrosen48/konfound
- https://github.com/data-edu/tidyLPA
- If you find yourself copying and pasting the same code, it may be worthwhile to encapsulate your code within a function (more on this in a moment!)

---

### B. Iteration

Iteration is another name for your computer carrying out some step(s) multiple times.

Iteration is helpful when, even after writing a function, you find yourself copying and pasting the same code (with modifications)

---

### B. Iteration

Another way of writing the above code in which we are scaling multiple variables:

```{r}
mpg %>% 
  mutate_at(vars(cty, hwy, cyl), standardize_variable) %>% 
  select(cty, hwy, cyl, everything()) # bring the transformed variables to the beginning of the data frame
```

---

### B. Iteration

Iteration can be helpful when you want to apply a function multiple times.

For example, I recently needed to download 14 surveys for a teacher professional development-focused
research project, https://megabitess.org!

There is a great package, qualtRics, which can help with this, but it's tedious to have to download the surveys one-by-one.

The map functions can help with this. 

---

### B. Iteration

The `purrr` package is a package that is part of the tidyverse and lets you do various iteration operations

The package consists of a family of related functions based on `map`

```{r, eval = FALSE}
library(qualtRics)
library(tidyverse)

# qualtRics::all_surveys()

description <- c("effectiveness - communication",
                 "effectiveness - summer",
                 "end-of-day survey",
                 "post-summer survey",
                 "pre-summer survey",
                 "end-of-day survey",
                 "end-of-day survey",
                 "end-of-day survey",
                 "end-of-day survey",
                 "post-summer survey",
                 "effectiveness - summer",
                 "effectiveness - GIS",
                 "post-GIS survey",
                 "pre-GIS survey"
)

survey_id <- c("SV_3IP9R9vFxZ7qGUe",
               "SV_bJEe2AWjZdH4dcp",
               "SV_81baWN1LtwNzV9r",
               "SV_5jue08g3DvyGlmd",
               "SV_6s9qu47PyM0js8t",
               "SV_5iNHeOZns6zzxat",
               "SV_9uhuHORliSFZq6x",
               "SV_d0EQx7RXOZxasT3",
               "SV_a46oo0Kklrlnh0F",
               "SV_e2PwPb0wmRAWCcl",
               "SV_9Zz8bjbBUPw1K4t",
               "SV_cFRH5pgKwnYJRLT",
               "SV_8HVSt4wfW3Hj2pn",
               "SV_6AoaxSCPgUlfNUV"
)
```

---

### B. Iteration

```{r, eval = F}
my_surveys <- tibble(description = description,
                     survey_id = survey_id)

all_surveys <- my_surveys$survey_id %>% map(fetch_survey)

my_surveys_end_of_day <- my_surveys %>% 
  filter(description == "end-of-day survey")

all_end_of_day_surveys <- my_surveys_end_of_day$survey_id %>% map_df(fetch_survey)
```

---

### C. Shiny continued: Reactivity

Shiny apps work on the principle of Reactivity:

- Functions in your server code need to run again when there are changes to the input
- UI is the same for all users of your app, server changes for each user
- The input object can only be modified in the UI, and can only be accessed in a reactive context
- We don't want to tell the app when to re-run code
- Shiny figures out when to run the code, you provide recipes for what to do if that happens
- When code is run depends on the connections between reactive dependencies, not the order written

---

### C. Shiny continued: Reactivity

Shiny apps work on the principle of Reactivity:

- Inputs from the UI connect to outputs in the server
- Inputs are modified based on user input, server code then runs reactively to change output accordingly
- Reactive expressions are a middle ground between inputs and outputs
- Reactive expressions change based on input, but they can be used in constructing outputs
- This can allow you to avoid repeating yourself
- This is necessary because using a reactive context lets your code respond to changing input values

---

### C. Shiny continued: Controlling when code is run

Sometimes you don't want your code to run right away

A tool that shiny provides to allow you to control this behavior is an action button 

An action button is a UI widget, but by using it in conjunction with the function `eventReactive` code running is dependent on the button, rather than reactive response to changing inputs

Observers allow you to make other reactive side effects dependent on actions e.g. `observeEvent`

Note: the most up to date version of Shiny recommends the use of a new function `bindEvent`

---

# Logistics

**This week**

* Homework 12: Available Thursday; **Due by Tuesday, 11/2**

* Final Project: More detailed plan present to class: Thursday (~ 3-5 minutes, ~ 2-3 slides)

---

## Schedule

- The product for your final project will be due are both due by the end of the day on Tuesday December 7
- We will do presentations of your final project in class on our last day - Tuesday November 30

---

### Wrapping up

On Slack:

- What is one thing you learned today?
- What is something you want to learn more about? 
- Share your feelings in GIF form!
