---
title: "Week 8 - Data Exploration and Summary" 
subtitle: ""
author: "Alex Lishinski"
date: "October 5, 2021"
output:
  xaringan::moon_reader:
    css: [theme.css]
    lib_dir: libs
    nature:
      highlightStyle: github
      highlightLines: true
      countIncrementalSlides: false
---

```{r, include = FALSE}
knitr::opts_chunk$set(echo = TRUE, fig.align = "center", message = FALSE)
```

# Welcome!

Welcome to *week 8*!

```{r echo = FALSE, out.width = "70%"}
```

**Record the meeting** 

```{r, include = TRUE, echo = FALSE}
library(here)
library(tidyverse)
```

---

### Review of last week's class

- Using color

- Grouping and stacking bar charts

- Faceting charts

---

### This week's topics

**Overview**

A. Summarizing variables

B. Inspecting data frames and correlations

C. Summary tables  

---

### A. Summarizing variables

**Overview**

* Sometimes you need to summarize variables

* Data viz is often a good tool to use

* Statistics are important too

As always with R, there are multiple ways to get there.

---

```{r, echo = FALSE}
library(emojifont)
```

### A. Summarizing variables - Base R

`library(`r emoji('man_shrugging')`)`

```{r}
summary(mtcars)
```

---

### A. Summarizing variables - skimr

skimr package - https://cran.r-project.org/web/packages/skimr/vignettes/skimr.html

```{r}
library(skimr)

skim(mtcars)
```

---

### A. Summarizing variables - skimr

Grouping also works with some of these summary functions

```{r}
mtcars %>% 
  group_by(am) %>%
  skim()
```

---

### A. Summarizing variables - psych

```{r}
library(psych)

describe(mtcars)
```


---

### A. Summarizing variables

Can also use subsets of variables

```{r}
mtcars %>% select(mpg, cyl, disp) %>%
  describe()
```

---

### A. Summarizing variables

Or subsets of rows

```{r}
mtcars %>% filter(str_detect(row.names(mtcars), "Merc")) %>%
  describe()
```

---

### B. Inspecting data frames and correlations

Another handy package is visdat: https://cran.r-project.org/web/packages/visdat/vignettes/using_visdat.html

Lets you essentially visualize the sort of information you see in `glimpse()`

```{r, out.width="50%"}
library(visdat)

vis_dat(storms)
```

---

### B. Inspecting data frames and correlations

Another useful function from visdat is `vis_compare()`

This lets you compare two data sets of the same size to see what's different

```{r, out.width="50%"}
storms_2 <- storms %>% mutate(day = as.numeric(day))

vis_compare(storms, storms_2)
```


---

### B. Inspecting data frames and correlations

Correlations are something you'll probably want to look at.

Base R:

```{r}
cor(mtcars)
```

---

### B. Inspecting data frames and correlations

Base R:

Correlations are in a matrix object

```{r}
storms %>%
  select_if(is.numeric) %>% 
  cor()
```

---

### B. Inspecting data frames and correlations

Base R:

Correlations are in a matrix object

```{r}
storms %>%
  select_if(is.numeric) %>% 
  cor(use = "pairwise")
```

---

### B. Inspecting data frames and correlations

corrr package - `correlate()`

Introduces a new data frame type for correlations

```{r}
library(corrr)

correlate(mtcars)
```

---

### B. Inspecting data frames and correlations

corrr package - Adds some useful features: variable subsets

```{r}
mtcars %>% correlate() %>%
 focus(mpg, cyl, disp)
```

---

### B. Inspecting data frames and correlations

corrr package - Adds some useful features: neater print output

```{r}
mtcars %>% correlate() %>%
 focus(mpg, cyl, disp) %>%
  fashion()
```

---

### B. Inspecting data frames and correlations

```{r, out.width="80%"}
mtcars %>% correlate() %>%
  rplot()
```

---

### B. Inspecting data frames and correlations

```{r, out.width="65%"}
mtcars %>% correlate() %>%
  shave() %>%
  rplot()
```

---

### B. Inspecting data frames and correlations

visdat package does correlation heatmaps as well:

```{r, out.width="65%"}
mtcars %>% vis_cor()
```

---

### C. Summary tables

We also can make some of this summary information neater for our Rmd documents

```{r}
describe(mtcars)
```

---

### C. Summary tables

We also can make some of this summary information neater for our Rmd documents

```{r}
library(knitr)
kable(describe(mtcars), digits = 2)
```

---

### C. Summary tables

We also can make some of this summary information neater for our Rmd documents

```{r}
cors <- mtcars %>% correlate() %>%
 focus(mpg, cyl, disp) %>%
  fashion()

cors

kable(cors)
```


---

### C. Summary tables

We can also do some additional customizing with `kable()`

```{r}
kable(describe(mtcars), digits = 2, caption = "This is my summary table")
```

---

### C. Summary tables

The kableExtra packages adds more customization options to `kable()`

```{r}
library(kableExtra)
table <- kable(describe(mtcars), digits = 2, caption = "This is my summary table")
row_spec(table, 5, bold = TRUE, background = "yellow") %>% 
  footnote("only the highlighted row is important, but you can't have a table with just one row")
```

---

### C. Summary tables

There are other packages that can customize tables for specific formats 

the sjPlot package is particularly useful for producing orderly and nice looking output tables for different kinds of models. 

Example: `lm()` summary output

```{r}
library(sjPlot)
model <- lm(mpg ~ disp + cyl, data = mtcars)
```


---

### C. Summary tables

Example: `lm()` print output

```{r}
model <- lm(mpg ~ disp + cyl, data = mtcars)
model
```

---

### C. Summary tables

Example: `lm()` summary output

```{r}
model <- lm(mpg ~ disp + cyl, data = mtcars)
summary(model)
```


---

### C. Summary tables

Example: `lm()` sjPlots table output

```{r}
model <- lm(mpg ~ disp + cyl, data = mtcars)
tab_model(model)
```

---

### C. Summary tables


```{r, eval = FALSE}
library(apaTables)
apa.cor.table(attitude)
apa.cor.table(attitude, filename="ex-CorTable1.doc")
```

---

# Logistics

**This week**

* Homework 7: Available thursday

* Readings
https://r4ds.had.co.nz/exploratory-data-analysis.html

* Schedule: How do we want to spend our time?

---

# Final Project

* [Final project]()

* [Slide deck to edit this week](https://docs.google.com/presentation/d/1KWU5bhxZmV63vkNtQNxHo-2YI_iYJmSW2oOZ_jtFeP0/edit?usp=sharing)
  * Must be logged in to UTK email/account
  
---

### Wrapping up

On Slack channel:

- What is one thing you learned today?
- What is something you want to learn more about? 
- Share your feelings in GIF form!

I really appreciate being able to see these reactions and get this feedback!
