---
title: "Homework 1"
author: "Alex Lishinski"
date: "8/25/2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, error = TRUE)
```

# Loading, setting up

## Checking that the project is correct

First, ensure that this `.Rmd` file is saved *in the directory for the project that you 
created for the course.* 

## First, let's install and load a package 

We'll do so by running the following line of code. This line of code, though, is 
a bit different from most: it installs a new R "package" with specific, additional 
functionality. Installing R packages is very common. The reason it is different from
most code, though, is you typically only want to run it once - after that, the package
is installed on your computer, and you don't need to run it again. In fact this line shouldn't even really be in this document, so if you already have the here package installed, you can just go ahead and delete everything from lines 27-29. Otherwise remove the # symbol, run the code, and then delete.

```{r}
# install.packages("tidyverse")
```

Next, we'll "load" the here package, to make it available for use in our current analysis / R 
session. Run this next chunk the same way as the first:

```{r}
library(tidyverse)
```

## Print a built-in dataset

Next, simply run msleep to view a dataset built-in to R:

```{r}
msleep
```

What do you notice? What do you wonder? Consider viewing additional rows by clicking
the numbers at the bottom of the table; or more rows by clicking the black right arrow 
at the top.

Add a few thoughts on what you notice and what you wonder after the dashes below!
You can add additional dashes, if needed!

- 
- 
- 



## Loading the tidyverse suite of packages

Next, load the tidyverse packages (or, more descriptively, the suite of packages)
Recall that the tidyverse includes a number of useful packages that work with each 
other and many other packages well; check out this site to learn more (I 
recommend poking around and familiarizing oneself with the tidyverse site!):

https://www.tidyverse.org/

Add your code below to load (hint: but not install, unless you were unable to 
do this in class) the tidyverse suite of packages. Remember to 
run the chunk by clicking the green arrow to the right of the chunk.

```{r}

```

We'll be using the gss_cat dataset for this homework. It is built-in to the 
tidyverse, so, to load it, just type its name in a code chunk and then run 
the code in that chunk.

## Examine data

Often, it is helpful and important to get to know the dimensions of your data.

How many rows and columns exist in the gss_cat data? Add the number after the two 
clauses below. It may help to add some code to determine this; you may find it helpful 
to add code on the line following this line, if you like.

```{r}
gss_cat
```

Number of rows:

Number of columns:

We have one more question to pose to you: What do rows and columns typically 
represent in your area of work and/or research? Rows typically represent "cases,"
the units that we measure, or the units on which we collect data. This is not a trick
question! What counts as a "case" (and therefore what is represented as a row)
varies by (and within) fields. There may be multiple types or levels of units
studied in your field; listing more than one is fine! Also, please consider what
columns - which usually represent variables - represent in your area of work and/or
research.

What rows typically represent:

What columns typically represent: 

# Selecting variables

Let's select only a few variables.

```{r}
gss_cat %>% 
  select(year, age, tvhours)
```

Let's *include one additional variable* in your select function.

First, we need to figure out what variables exist in our dataset (or be reminded 
of this - it's very common in R to be continually checking and inspecting your data)! 

We can do this a number of ways, including simply running the name of the dataset and looking at the printed output. There's also a function called `colnames()` that can be very handy.

```{r}
colnames(gss_cat)
```

Then, add a new variable to the code below, being careful to type the new variable 
name exactly as it appears in the data. We've added some code to get you started.

```{r}
gss_cat %>% 
  select()
```

# Filtering variables

Next, let's explore filtering variables. Check out and run the next chunk of code:

```{r}
gss_cat %>% 
  filter(tvhours > 3)
```

Note that *a bit* more information on this dataset is available by running ?gss_cat in 
the console tab of RStudio; or, by including ?gss_cat in a code chunk and then running
the code in that chunk. *A lot* more information on the large-scale survey from which
this data was accessed is here: https://gss.norc.org/get-documentation

What do you notice (about the code or its output)? What do you wonder? Add these
thoughts below!

- 
- 
- 

What do you think will happen if we add another condition to the filter statement?

**Try it out!**

Let's try to filter on age, as well. We've added some code to get you started.

Logical operators define relationships between conditions that are either TRUE
or FALSE. R calls variables with a TRUE or FALSE value "logical," but they're
otherwise known as "boolean." When you use the filter() function for example
you're defining a logical condition. You give the condition to filter on based
on variables in the data, such as `mpg < 20` or `am == 1`, and you are asking the
filter() function to return the rows for which that condition is TRUE, i.e. that
the mpg value on that row is less than 20 or the am value on that row equals 1.
Other logical operators can connect 2 conditions. You could use `mpg <20 & am
== 1` to filter the rows meeting both conditions, and you can use `mpg < 20 | am
== 1` to filter rows meeting one or more of the two conditions.

```{}
# equivalent to
==
  
# not equivalenet to
!=
  
# greater than
>
  
# lesser than
<

# equivalent or greater to
>=
  
# equivalent or lesser than
<=
```

Add a logical operator (of your choosing) and complete the logical statement for
the age variable below; you can use any of the logical operators defined above:

```{r}
gss_cat %>% 
  filter(tvhours > 3,
         age )
```

## Arrange

Let's arrange the data frame based on the tvhours variable:

```{r}
gss_cat %>% 
  arrange(tvhours)
```

**Try it out!**

Now, arrange the dataset based not on tvhours, but on a variable of your choice.
We've added a chunk of code below, but nothing more:

```{r}

```

## Reach 1

Can you compose a series of functions that include the select(), filter(), and
arrange functions? Recall that you can "pipe" the output from one function to the 
next.


```{r}
gss_cat %>% 
```

When content with the result, write a brief (1-2 sentence) interpretation of what 
your code outputs next:


## Reach 2

Can you read a .csv file using the readr package, HINT: we talked about it in the slides

There is a .csv file called answer_export.csv, that you can find here: 


Download it and assign it to the name data:

```{r}
library(readr)



exists("data")
```


## fin

Nice work! Once you've finished your work, click "Knit" at the top to render a 
report that you can share; this is how you'll submit homework for class! 

Please submit:

- Submit the .html file you used to render a report to Canvas
- Upload the .Rmd file you rendered to the #homework-yourname channel in Slack.

:tada: :tada: :tada: 

## Self-assessment and reflection

Respond to the following three questions on a 1 (not at all) to 5 (very much) 
scale by replacing the "x" below with your response:

```{r}
x = NULL
tibble::tribble(
  ~question,                                   ~response,
  "How challenging was this homework?",        x,
  "How interesting was this homework to you?", x,
  "How valuable was this homework to you?",    x
)
```

Include any other comments, feedback, or reflections on this homework below:



If you like, you can post these other comments, feedback, or reflections in the 
message you post to #homework in Slack to share what you've done.