---
title: "HW Week 4 - Advanced Data Cleaning Skills"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Background 

For this homework, we are going to be working with the Experience Sampling
Method - ESM - data that we used in class.

Here's a bit of an overview and a reminder of this dataset. The database
contains responses, corresponding question numbers, unique identifiers for the
survey and for the item level responses, and the course assignment that each
survey corresponds to. What it does not have are unique identifiers of the
person who sent it, i.e.:

- their phone number 
- the date and time when it was sent
- and, if we wanted later, other information, such as information about their major

That's where the Twilio (ESM) data file comes in it does have that information,
and importantly, it also uses the same unique identifiers for the ESM responses. 

It also has some junk in it that we *don't* want, like the messages
where students initiated but didn't take the survey.

To provide just a bit more context, students answered our five ESM survey items \~11 times
each, with some doing so as many as 18 times. 

So, it might be interesting to look at the five survey questions, and see what 
 average responses were for each question for each student.

In class we went over getting the mean by survey question overall, but this time we
want to know how each person responded on average to each item. There's \~75
people in this dataset, and, again five survey items.

**So, just to sum up, we have around 75 students each responding around 11 times 
to five questions. What we'd like to get is a summary data table that's 75 rows 
and 5 columns, with each entry being each person's average response for each of the 
five questions.** 

There's a lot here, and if you have questions about any of this, please message 
over in the #questions channel on Slack. You'll note that this is entirely data 
tidying/processing, but is very much the kind of thing that you may wish to do as 
the descriptive part of an analysis for a research project.

This task will involve a number of steps that we'll tackle one at a time, then at the
end we can put everything together and you will (maybe?) marvel at what you've done.

## Downloading the data and saving them to the data directory

The first thing you will need to do is get the data. You'll need the two ESM data
files: 

- `twilio_data.csv` 
- `database_data.csv`. 

Both can be acquired from the GitHub repository here:

https://github.com/alishinski/course-website/tree/master/content/data

To save them, simply run the following code chunk. But, first, load two packages:

- tidyverse
- here

Do this - load these two packages - in the next chunk:

```{r}

```

Then, let's download the files next. If you don't have a data subfolder, delete that part of the code.

```{r}
download.file("https://github.com/alishinski/course-website/tree/master/content/data/twilio_data.csv",
              destfile = here("data", "twilio_data.csv"))

download.file("https://github.com/alishinski/course-website/tree/master/content/data/database_data.csv",
              destfile = here("data", "database_data.csv"))
```

Let's break down what's going on:

- First, we load the here package so that we can specify where we want to save the file
- Second, we use the `download.file` function -- twice. First, we download the 
`twilio_data.csv` file; second, we download the `database_data.csv` file

If you saw two message with "trying URL" and then the URL and some information 
about the file and the word "downloaded", you should be good to go! I recommend
checking your project directory to make sure that both files are there.

## Reading the data

Now, you'll need to read the data. We'll do this for the first file; you do this 
for the second file. 

```{r}

```

## Joining the data

The next thing we need to do is join the two data frames since we need
variables from both. We only want to keep the matching rows from both data sets. 

At each intermediate step in the process, starting now, you should assign your 
results to a data frame. We'll build this up iteratively, so each step will involve
chaining the new step to what you did in the previous step.

First, here, join the two data frames together based on a common variable. Remember
the differences between the different types of joins! More on those here:

https://making-data-science-count.github.io/s21-intro-to-data-sci-methods-in-ed/slides/presentations-week-4.html#1

You'll need one of the join functions; you may find this help documentation helpful:
https://dplyr.tidyverse.org/reference/join.html

```{r}

```

Now that the data is joined, we will want to get rid of some of the variables
that we won't need. 

So, consider what our end goal is: We want a table of
means by person and by item. 

- We will need the answers to obtain these means, so we'll
need the content variable; 
- We will need to know which question they go with, so we'll need the question ID; 
- We will also need to know which person each response goes with, so that's the From variable,
with the anonymized phone numbers; 
- Besides these, it would also be helpful to have a unique identifier so that we could rejoin this data with the main set later; so we should keep the survey ID
variable. 

In this next step let's pare down our data set to just those important
variables:

```{r}

```

Now that we've gotten rid of the unimportant variables, that will make our next
steps a bit simpler. There's more than one way to go from here, but let's think
about our end goal again. 

We want means by item (and also by person), so one way to do this would involve 
separating the content variable by item, so that each item
has it's own variable. If we can do that, we can then group by person,
summarize, and achieve our desired result. 

Separating the content variable by item is exactly what the `pivot_wider()` 
function can help us do. This one can be tricky, so we'll fill in some parts for
you to get started, but you'll need to add the crucial bits, and connect it with 
what you've done before.

```{r}

```

Great! Now that you have those new variables for each question, all that's
left to do is to group the data and then summarize on each of those five variables.
With the summarize function, you can do several summaries, so it's just a matter
of doing it for each one. Let's do that here.

```{r}

```

So now you might be mad, because that didn't work and it gave you an obnoxiously
long error message. Sorry about that. But maybe you also remember that we
covered dealing with this exact issue in the lecture. And let's also drill down
on what this means `argument is not numeric or logical: returning NA` you can't
take the mean of something that's not a numeric vector (or logical vector, but
don't worry about the logical part right now). So we'll need to fix that, but
you already know how from the lecture and last week's homework, so go ahead and
do that below.

```{r}

```

Now that you have everything, just type in the name of your summary data table so 
we can see the print output.

```{r}

```

If you made it all the way, congrats, you did a great job! If you couldn't get
it quite 100% right, don't worry, you're probably at least 80-90% there, and
we'll help you the rest of the way.

## fin

You know the drill; render and please submit:

- Submit the .html file you used to render a report to Canvas
- Upload the .Rmd file you rendered to the #homework channel in Slack.

One difference from last week: Please post a comment, feedback, or reflection
below *as well as to Slack when you post your homework*.

## Self-assessment and reflection

Respond to the following three questions on a 1 (not at all) to 5 (very much) 
scale by replacing the "x" below with your response:

```{r, reflection}
x = NULL
tibble::tribble(
  ~question,                                   ~response,
  "How challenging was this homework?",        x,
  "How interesting was this homework to you?", x,
  "How valuable was this homework to you?",    x
)
```

Include any other comments, feedback, or reflections on this homework below:


Important note: Please post your comment(s), feedback, or reflections in Slack when you share your Rmd file!