---
title: "HW Week 5 - Introduction to Data Viz"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, eval=FALSE)
```

BEFORE YOU START, FIRST REMOVE THE `, eval=FALSE` FROM THE CODE ABOVE

# Background 

We will use a new dataset for this week's homework, the tidykids dataset. 

You can read more about it here: https://jrosen48.github.io/tidykids/

# Loading data

To download the data, run the following code chunk. Like for last week, this function
uses two arguments:

1. The URL on the Internet for the file
2. `destfile`, where the file will be saved on your computer

Note that - like for last week - you will need to have created a `data` directory in
the same directory in which this file is saved for this to work.

```{r}
download.file("https://raw.githubusercontent.com/making-data-science-count/s21-intro-to-data-sci-methods-in-ed/main/data/tidykids.csv", destfile = here("data", "tidykids.csv"))
```

A brief aside: The above function has two arguments, but that *only one argument*, destfile, 
is named. The first argument, which is named url, is not named because, by convention, 
the first arguments to functions are not named. To see the arguments for this and any 
other function, add a question mark before the name of the function and run that line of code in 
the console, e.g.: 

?download.file

Next, let's read the tidykids data using the read_csv() function. Save the data
to the name tidykids.

```{r}

```

Finally, in the code chunk below, show how you would inspect the data you just loaded - 
does it appear as you think it should appear? What do you notice from taking a first 
look at your data?

```{r}

```

A simple codebook for the data, which is state-level data related to spending on 
services for children, is available here:

https://jrosen48.github.io/tidykids/reference/tidykids.html

We will take one last step before proceeding to visualizing the data; for this
homework, filter the variable column to only include a single variable, such as,
for example, PK12ed or pell. There are many more; a 
description of the different values within the variable column is available here:

https://jrosen48.github.io/tidykids/articles/tidykids-codebook.html

You will need to add the name of the variable you select (choose whichever you 
find the most interesting)! in the next code chunk; we've written some
code to get you started:

```{r}
tidykids_variable <- filter()
```

Next, let's take one more look at our data, just to make sure that the above code
worked as we intended. Write that code to inspect or check out your data in the 
next chunk:

```{r}

```

# Overview

The remainder of this homework is focused around three visualization-related tasks:

1. Visualizing one variable
2. Visualizing two variables
3. Theming and fine-tuning a plot

There should be six columns in your dataset. If you have not yet, take a careful
look at the codebook for the dataset to understand what each of the six columns 
represent: 

https://jrosen48.github.io/tidykids/reference/tidykids.html

# Visualizing one variable

The following are some "geoms", or geometric shapes, with which you can create visualizations
of one variable:

- geom_histogram()
- geom_density()
- geom_freqpoly()
- geom_bar()

The code below creates a plot - but one without any geometric shapes, so it's not very 
interesting (you can run the code to see what we mean!). To this code below, add
one of the above geoms. Recall that we add layers to ggplot2s (so, any line
after the line on which we call ggplot()) through a different operator than we otherwise
use in R (the pipe operator). Instead, we use the + operator. Take a look [at the 
slides for this week](https://making-data-science-count.github.io/s21-intro-to-data-sci-methods-in-ed/presentations.html) if you need a pointer on this.

```{r}
tidykids_variable %>% 
  ggplot(aes(x = inf_adj_perchild))
```

If you were successful, you should see something much more interesting than what appeared before;
you should see a plot that summarizes the spread, or the distribution, of values.

In the next code chunk, run the same code, but change the variable that you specify for 
the x aesthetic.

```{r}

```

Next, select a different geom than the one you selected; choose which variable
you wish to represent (by specifying it as mapping to the x aesthetic), too:

```{r}

```

Finally - for this portion of the homework - choose one more geom
(other than one you have already used), and create a plot using that geom below:

```{r}

```

What do you notice? What do you wonder? At at least one reflection below (but feel free
to add additional dashes to add additional reflections!)

- 

There is one last task pertaining to visualizing a single variable. Next, filter
the dataset once more, this time to include data from only a single year. You 
choose which year! Note the type of the year variable; you might wish to treat it 
as a character or as an integer; refer to homework 3 for a reminder for how to change
the type of a variable. We strongly recommend saving the result of your filter to a 
data frame with a different name than those that are already used.

Once you have filtered the data, create one additional plot (including any of the
elements utilized in the plots above) in the chunk below:

```{r}

```

# Visualizing two variables

Visualizing two variables invites a number of additional considerations. We will
focus on only one geom for now, adding others over the next few weeks:

- geom_col()

For the next visualization (on visualizing two variables), use the
dataset you filtered to include only responses from a single year.

Using the below as a starting point, specify the state variable as either the x
or y aesthetic, and specify another variable as the other (depending on which you
choose for state):

```{r}
d %>% 
  ggplot(aes()) +
  geom_col()
```

You might see something that looks interesting, but difficult to interpret.

The reorder() function can be very helpful for making such plots more interpretable.

Here is an example (using code that is not run - note the code chunk option 
eval = FALSE) of how the reorder function can be used:

```{r, eval = FALSE}
my_data %>% 
  ggplot(aes(x = reorder(class_level, mean_value), y = mean_value)) +
  geom_point()
```

The above code is not quite useful to you for at least two reasons:

1. The data and variables are named differently than those you are using
2. We are using a different geom

But, using the above as a starting point, can you reorder the columns in the
first visualization that you made using geom_col in the code chunk below:

```{r}

```

There is a *lot* more you can and will do, including visualizing patterns over time
and creating "small multiples", or separate plots for each year (or, even, for each variable).

# Theming and fine-tuning a plot

There are many ways to theme and fine-tune a plot. For now, we'll focus on two 
changes you can make, those enabled through:

1. The addition of a theme, which changes many parts of a plot
2. The modification of a specific aspect of a plot (specifically, its theme, even 
if that is the default theme)

In the next chunk, copy the code from the last code chunk you ran. Then, add
another layer (as you have done to add anything to your ggplot2 plots - with the 
+ operator) with one of the built-in themes. Here is a brief list; you can also see
additional options by typing ggplot2::theme_ in the console (or in a code chunk)
and then noting the names of themes that autocomplete:

- theme_bw()
- theme_minimal()
- theme_dark()

Add a theme to your plot below:

```{r}

```

Next, as one example of many ways that you can fine-tune your plot, copy and paste
the code you just ran in the chunk above into the chunk below, and change the x-axis 
label. You can do so by adding xlab() as a layer, with the name of the label as 
you wish it to appear as the sole argument to the xlab function:

```{r}

```

# Reach 1

Are the built-in options not very exciting to you?

Install the ggthemes package and try out one of the themes include in it along 
with code for a plot in the chunk below; see the themes here:

https://ggplot2-book.org/polishing.html

```{r}

```

# Reach 2

Can you create a visualization that represents differences by year? To do so, make
sure you use the data frame that contains all of the years. You may need to process,
tidy, or summarize the data to create the visualization you have in mind; try this
below.

```{r}

```

## fin

Render and submit:

- Submit the .html file you used to render a report to Canvas
- Upload the .Rmd file you rendered to the #homework channel in Slack.

One difference from last week: Please post a comment, feedback, or reflection
below *as well as to Slack when you post your homework*.

## Self-assessment and reflection

Respond to the following three questions on a 1 (not at all) to 5 (very much) 
scale by replacing the "x" below with your response:

```{r, reflection}
x = NULL
tibble::tribble(
  ~question,                                   ~response,
  "How challenging was this homework?",        x,
  "How interesting was this homework to you?", x,
  "How valuable was this homework to you?",    x
)
```

Include any other comments, feedback, or reflections on this homework below:


Important note: Please post your comment(s), feedback, or reflections in Slack when you share your Rmd file!