---
title: "HW Week 6 - Introduction to Data Viz"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Background 

We will use a built in dataset for this week's homework, called `mpg`. It
contains fuel economy and other data on 38 models of cars. You can learn more
about it by typing this in the console `?mpg`.

# Loading data

To load the data, run the following code chunk. 

```{r}
library(tidyverse)

data(mpg)
```

Once you've loaded it, familiarize yourself with the different variables in the
dataset using the methods that we've talked about.

# Using the `count()` function

As we discussed in class, the `count()` function can be a useful way to
summarize data for plotting. Here's an example of how this works:

```{r}
mpg %>% count(year)
```

Below, we would like you to create a barplot showing the number of vehicles per
manufacturer, using the `count()` function and `geom_col()`.

```{r}


```

# Plotting summary statistics

As we discussed in class, the combination of `group_by()` and `summarize()` can
be used to generate summary dataframes which you can then use to plot summary
statistics. Below, we would like you to generate a summary dataframe and plot
the summary statistics you create. You can use whatever variables from the data
you see fit. Refer back to the slides and video from this week's (and last
week's) class for examples of how to do this if you're not sure.

```{r}

```

# Recoding a variable for plotting

In class this week we showed you examples in the slides and in the live coding
demo of how to recode a variable for plotting purposes. Below, we would like you
to use the `recode()` function from dplyr to recode a variable from `mpg` to
make a plot. One natural way to implement this would be to condense the number
of categories in one of the above variables to make a more manageable number to
show in a plot. However you are free to be creative and implement this as you
see fit.

```{r}

```

# Tell us a story

As we mentioned in one class, one important principle of creating data
visualizations is using them to tell a clear story about some aspect of the
data. We would like you to tell us something interesting about the data. By
interesting we don't mean it has to be anything earth-shattering, but just a
coherent conclusion about the data that is communicated by your plot. Make it
look nice insofar as it serves this end, but there's no need to go overboard.
For more about what we mean by tell a story, we will refer you to this book
chapter which shows some examples of storytelling with the mpg dataset:
https://r4ds.had.co.nz/graphics-for-communication.html
Of course we would like your story to be different from the examples in that
chapter. Be creative and have fun with it!

```{r}

```

# Reach 1

*Note: The reaches are not required to complete, but are chances for you to do more if you wish to*

The first task in the homework asked you to plot the count of different vehicles
by manufacturer. This should have been relatively straightforward to plot the
number of different vehicles from each manufacturer, but many of these are the
same model. Below, try to produce a plot of the number of unique models for each
manufacturer. For example, there are 3 unique Audi models in the dataset: a4, a4
quattro, and a6 quattro. HINT: You can supply more than one variable to
`count()`, and you'll need to use the `count()` function twice to get the right
counts.

```{r}

```

# Reach 2

Take the plot you made above, your "story" and make it look extra pretty. Change
2 or more aspects of the theming of your plot to make it more "publication
quality" whatever that means to you.

```{r}

```


## fin

Render and submit:
- Submit the .html file you used to render a report to Canvas - Upload the .Rmd
file you rendered to the #homework channel in Slack.
One difference from last week: Please post a comment, feedback, or reflection
below *as well as to Slack when you post your homework*.

## Self-assessment and reflection

Respond to the following three questions on a 1 (not at all) to 5 (very much)
scale by replacing the "x" below with your response:

```{r, reflection}
x = NULL
tibble::tribble(
  ~question,                                   ~response,
  "How challenging was this homework?",        x,
  "How interesting was this homework to you?", x,
  "How valuable was this homework to you?",    x
)
```

Include any other comments, feedback, or reflections on this homework below:


Important note: Please post your comment(s), feedback, or reflections in Slack
when you share your Rmd file!