---
title: "HW5-6"
output: html_document
---

# Background

For this homework, we will be covering material from the last 2 weeks of data visualization.

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Background Part 1

We will use a built in dataset for part 1 this week's homework, called `mpg`. It
contains fuel economy and other data on 38 models of cars. You can learn more
about it by typing this in the console `?mpg`.

# Loading data

To load the data, run the following code chunk. 

```{r}
library(tidyverse)

data(mpg)
```

Once you've loaded it, familiarize yourself with the different variables in the
dataset using the methods that we've talked about.

# Using the `count()` function

The `count()` function can be a useful way to
summarize data for plotting. Here's an example of how this works:

```{r}
mpg %>% count(year)
```

Below, I would like you to create a barplot showing the number of vehicles per
manufacturer, using the `count()` function and `geom_col()`.

```{r}


```

# Plotting summary statistics

As we discussed in class, the combination of `group_by()` and `summarize()` can
be used to generate summary dataframes which you can then use to plot summary
statistics. Below, we would like you to generate a summary dataframe and plot
the summary statistics you create. You can use whatever variables from the data
you see fit. Refer back to the slides and video from this week's (and last
week's) class for examples of how to do this if you're not sure.

```{r}

```

# Tell a story

As I mentioned in one class, one important principle of creating data
visualizations is using them to tell a clear story about some aspect of the
data. I would like you to tell me something interesting about the data. By
interesting I don't mean it has to be anything earth-shattering, but just a
coherent conclusion about the data that is communicated by your plot. Make it
look nice insofar as it serves this end, but there's no need to go overboard.
For more about what we mean by tell a story, I will refer you to one of the readings 
for this week, which shows some examples of storytelling with the mpg dataset:
https://r4ds.had.co.nz/graphics-for-communication.html
Of course I would like your story to be different from the examples in that
chapter. Be creative and have fun with it!

```{r}

```

# Background Part 2

We will use a data set built into dplyr for part 2 of this week's homework. 
The data, which we have encountered before,
"includes the positions and attributes of 198 tropical storms, measured every six 
hours during the lifetime of a storm." 

Like for the `mpg` data set, you can learn more about it by typing 
this in the console `?storms`. Note that the back tick marks are simply to format
code in R Markdown - you do not include these when you run the code. 

Note how this appears differently in output; in fact, since Slack allows us to
use Markdown, you can also use back ticks like this to format code in Slack!).

# Loading data

To load the data, in the next code chunk, load the tidyverse package and then type
storms to load and view the storms data from the dplyr package that is loaded as a 
part of the tidyverse:

```{r}


```

Once you've loaded it, familiarize yourself with the different variables in the
data set using the methods for inspecting and exploring a new data set we have 
discussed.

Next, filter the data to include observations only from 2006 (including that year)
through the last year with data available in the data, and save the results back
to a new object to use for the remainder of this homework.

```{r}


```

# Using Color

First, create a plot using pressure and wind on the x- and y- axes (which axis 
corresponds to which is you choice)! Use geom_point() in this plot.

```{r}

```

**Assigning a color to a geom**

Next, assign a color to the points by adding to the code above.

To see a list of available colors in R, type `colors()` in the console. You can also consult this resource: http://www.stat.columbia.edu/~tzheng/files/Rcolor.pdf?utm_source=twitterfeed&utm_medium=twitter

If helpful, check out this last week's slides for a tip: 
https://f21-intro-to-data-sci-methods-in-ed.netlify.app/slides/week7-slides.html

```{r}

```

**Mapping a color scale to a variable**

Next, map a color scale to a variable. This is different from assigning a color
to points; instead, the use of color should be used to represent different values
of a variable. 

Choose which variable you would like to represent with a color scale. For a tip,
consult this week's slides (linked above).

```{r}


```

**Changing the color scale**

Next, change the color scale. Recall the function we demonstrated in our slides
for easily changing the color scale (hint: it includes the term "brewer"). In your
use of this function, consider whether the specific color scale you are using is:

- sequential,
- diverging, or
- qualitative

And, specify the correct scale for the variable you are representing corresponds 
to the variable you wish to represent. Finally, once you have settled on the type of 
color scale you are using, consider changing the color palette; view the help documentation 
(or the slides) for tips on how to do that. Feel free to play around with what 
variable you represent, what scale you use, and what palette you use!

```{r}

```

# Stacking and dodging bars

By default, the geom we use for bar/column charts, `geom_col()`, "stacks" the bars 
vertically. 

First, run the following code:

```{r}
storms %>% 
  group_by(year) %>% 
  summarize(mean_wind = mean(wind))
```

What do you notice? Add a note about below:

- 

Next, replace year in the code above with status, and run the new code:

```{r}

```

What do you notice now? Add a note about this below:

- 

Finally, group by both year and status; add both to your `group_by` function.

If you need a tip regarding how to group by multiple variables, check the code 
used in some of the slides, or the help documentation (by running `?group_by()`
in the console.

```{r}

```

What do you notice now? Add one last note:

- 

**Plotting the mean wind speeds**

Next, plot the mean wind speeds, using the code you wrote above as the 
foundation. Specifically, using ggplot2, assign the x-axis values to the year,
and the y-axis values to the variable you calculated for the mean wind speed. 
Finally, specify that the fill aesthetic is equal to the location variable; see 
the slides for an example of using the fill aesthetic.

```{r}

```

Finally, dodge the plots by year. To do so, add an argument to dodge the columns
to the `geom_col()` function you used; our slides have an example that might help
you to do this:

```{r}

```

# Faceting plots

Above, beneath the bold text **Changing the color scale**, you created a scatter plot
that involved mapping a color to a variable and changing the scale to be appropriate
for the variable you chose. 

Below, copy and paste the final code you used. Add to that code a layer that will
facet the plots based upon the year. An example of how to do this is available in 
the section of the slides corresponding to faceting plots. Depending on what variable
you mapped to a color, you may need to tinker with your plot (and the variables you 
included in it) for this plot to make sense. Tinkering and modifications, including
to theme elements, are welcome.

```{r}

```

## fin

Render and submit:
- Submit the .html file you used to render a report to Canvas - Upload the .Rmd
file you rendered to the #homework channel in Slack.
One difference from last week: Please post a comment, feedback, or reflection
below *as well as to Slack when you post your homework*.

## Self-assessment and reflection

Respond to the following three questions on a 1 (not at all) to 5 (very much)
scale by replacing the "x" below with your response:

```{r, reflection}
x = NULL
tibble::tribble(
  ~question,                                   ~response,
  "How challenging was this homework?",        x,
  "How interesting was this homework to you?", x,
  "How valuable was this homework to you?",    x
)
```

Include any other comments, feedback, or reflections on this homework below:


Important note: Please post your comment(s), feedback, or reflections in Slack
when you share your Rmd file!