---
title: "Homework Week 8 - Data Exploration and Summary"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Background 

This week we will use a real data set of course data from 614 undergraduate
students. The data consists of students' scores on different course assignments,
as well as several background variables collected for research purposes.

The data set can be downloaded from the github repo here:
<https://github.com/alishinski/course-website/blob/master/content/data/undergrad_data.csv>

The data consists of an identifier, columns on project scores, a total course
points column, columns on exam scores, scores on a pre and post problem solving
ability test, 5 personality trait variables, and 4 motivation variables.

# Loading data

Download the .csv file and then load it in this chunk. But first, load the
packages you will need (you can always come back and add more):

```{r}


```

Once you've loaded it, familiarize yourself with the different variables in the
data set using the methods for inspecting and exploring a new data set we have 
discussed, particularly using the `visdat` package. 

```{r}


```

What are some observations you made about the data from your exploration?


# Data Report 

We would now like you to use the undergrad data to produce a report including
the following elements:

# Summary table

Make a nicely styled summary table using the `kable()` function from the `knitr`
package, using whichever summary function you want (check the slides from this
week to see the different options we discussed). But first be sure to exclude
any variables where it doesn't make sense to have a numeric summary, e.g.
character variables.

```{r}

```

Now please use the `kableExtra` package to make some additional formatting
tweaks. highlight the lowest and highest values of some column of your choice
using the `row_spec()` or `cell_spec()` functions, and then make one other
formatting addition of your choice, like a title or a footnote or whatever you
want.

```{r}

```

# Correlation Matrix

Make a correlation matrix, including only the 11 individual project score
variables, and style that output nicely. 

```{r}

```

Now use the `corrr` package to produce a correlation heat map for the
correlation matrix you just made, and use the lower triangle only.

```{r}

```

# Regression model

Run a regression model using `lm()` with the `Total_Pts` variable as the
dependent variable, plus whichever independent variables you want. If you have
questions about how to use `lm()` to specify models, you can search (the web)
for "lm syntax r"; a number of good tutorials and resources will appear, and
we can recommend a few if you would like to ask us for recommendations.

```{r}

```

Now use the `sjPlot` package to create a nice output table for your `lm()` model
as we showed in class. 

```{r}

```

Now we have one last task for you that may be a bit harder. We would like you to
create a composite variable by creating a sum of the three exam score variables in
the data set.

We didn't cover this in class, but here's a Stack Overflow post where a few
different answers using different methods are provided:
<https://stackoverflow.com/questions/10945703/calculate-row-means-on-subset-of-columns>

Now make a new `lm()` model, using this composite variable as the dependent
variable, and generate the output table using the `sjPlot` package just as in
the last step.

```{r}

```

Remake the correlation matrix from above using only the values that are not
equal to 0. Also generate the significance values of the correlations.

<http://www.sthda.com/english/wiki/correlation-matrix-a-quick-start-guide-to-analyze-format-and-visualize-a-correlation-matrix-using-r-software#correlation-matrix-with-significance-levels-p-value>

```{r}

```

## fin

Render and submit:
- Submit the .html file you used to render a report to Canvas - Upload the .Rmd
file you rendered to the #homework channel in Slack.
One difference from last week: Please post a comment, feedback, or reflection
below *as well as to Slack when you post your homework*.

## Self-assessment and reflection

Respond to the following three questions on a 1 (not at all) to 5 (very much)
scale by replacing the "x" below with your response:

```{r, reflection}
x = NULL
tibble::tribble(
  ~question,                                   ~response,
  "How challenging was this homework?",        x,
  "How interesting was this homework to you?", x,
  "How valuable was this homework to you?",    x
)
```

Include any other comments, feedback, or reflections on this homework below:


Important note: Please post your comment(s), feedback, or reflections in Slack
when you share your Rmd file!
